---
title: "SPCN V506 Homework Exercise 4"
author: "Author(s): Shahapurkar, Gangaprasad, Logan Mcguire, Hadi Abbas"
date: "`r Sys.Date()`"
output: 
  word_document:
    toc: true
    toc_depth: 3
    number_section: false
knit: (function(input, ...) {
  rmarkdown::render(
    input,
    output_file = paste0(
      xfun::sans_ext(basename(input)), '-', Sys.Date(), '.doc'
    ),
    output_dir = here::here("output"),
    envir = globalenv()
    )
  })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      eval = TRUE, 
                      message = FALSE,
                      tidy = TRUE)
```

```{r library, echo=FALSE, warning=FALSE}
if(!require('pacman')) {
  install.packages('pacman')
  library('pacman')
  pacman::p_load(pacman, Rmisc, dplyr, DescTools, car, ggplot2, psych, sjstats, 
                 rmarkdown, freqdist, readxl, foreign, here, formatR)
}
```
## **Part II: - R Programming, Inferential Statistics, and Data Visualization** 


```{r}
root_dir = here::here()
```

```{r}
fileRealEstate = file.path(root_dir,"data","real_estate.csv")
real_estate <- read.csv(fileRealEstate)

fileschools = file.path(root_dir,"data","schools.csv")
schools <- read.csv(fileschools)

fileblsa = file.path(root_dir,"data","blsa.csv")
blsa <- read.csv(fileblsa)
```

### **Problem 12:**

$\mu_{Ocearn}$ = Mean number of cars in Ocean drive parking lot

$\mu_{Rio}$ = Mean number of cars in Rio Rancho drive parking lot

$H_{0}:\mu_{Ocearn} = \mu_{Rio}$ and $H_{1}:\mu_{Ocearn} \ne \mu_{Rio}$

Two-tailed test.

$\alpha = 0.05$ and Degrees of Freedom = 51 Critical t-value range is -2.008 < t < 2.008


```{r}
print(qt(p=.05, df=51, lower.tail=TRUE))
```



```{r}

ocean_drive = c(89,115,93,79,113,77,51,75,118,105,106,91,54,63,121,53,81,115,67,53,69,95,121,88,64)
rio_rancho = c(128,110,81,126,82,114,93,40,94,45,84,71,74,92,66,69,100,114,113,107,62,77,80,107,90,129,105,124)
test_result = t.test(ocean_drive, rio_rancho, conf.level = .90, var.equal = FALSE);test_result
```

Computed p-value =  0.3795


Since the p-value is larger than the significance level, we fail to reject the null hypothesis.

The computed t-value is within the critical range so we fail to reject the null hypothesis. There is no significant difference in the mean number of cars.

### **Problem 13:**

```{r problem13}
c1 = c('Holt', 'Pierse', 'Merenick', 'Lanoue', 'Fagan', 'Bobko', 'Kippert', 
       'San Roman', 'Kurimsky', 'Davison', 'Lozier', 'Cieslinski', 'Rowatti', 
       'Koppel', 'Rumsey', 'McGinnis', 'Pierce', 'Roll', 'Lang', 'Miller')
c2 = c(17, 20, 29, 43, 36, 43, 45, 19, 49, 49, 35, 16, 23, 33, 44, 44, 28, 29, 39, 22)
c3 = c(10, 39, 37, 27, 12, 41, 24, 26, 28, 26, 32, 32, 21, 12, 40, 42, 22, 19, 35, 12)
income = data.frame(
  Homeowner = c1,
  Home_5yr_Ago = c2,
  Home_Now = c3
)
print(head(income))
t.test(income$Home_Now, income$Home_5yr_Ago, alternative = "less")

```
We reject the null.


### **Problem 14:**

```{r problem14_1}
obs_freq = c(140, 120, 90, 50)
exp_freq = c(100, 100, 100, 100)
p_value = exp_freq / sum(obs_freq)
chisq.test(obs_freq, p=p_value )
```
Reject the null

### **Problem 15:**


$H_0:$ There is no difference in the proportion of Isaac employees visiting the doctor

$H_1:$ There is a difference in the proportion of Isaac employees visiting the doctor

We selected the .01 significance level. The probability is .01 that a true null hypothesis is rejected.

The test statistic follows the chi-square distribution.

The number of degrees of freedom is k - 1 =  4 - 1 = 3. The critical value for 3 degrees of freedom and the .01 level of significance is 11.345. 
*

```{r problem15}
visits <- c(29, 11, 16, 4)
efpca <- c(.53, .19, .17, .11)
chisq.test(visits, p=efpca)
```
The computed $X^2$ of 4.58 is less than the critical value. The decision therefore is not to reject null hypothesis. The difference between the visits could be by chance.

### **Problem 16:**

$H_0:$ There is no relationship between age and amount of pressure

$H_1:$ There is relationship between age and amount of pressure

We selected the .01 significance level. The probability is .01 that a true null hypothesis is rejected.

The test statistic follows the chi-square distribution.

The number of degrees of freedom is $df = (r-1)(c-1) = (4-1)(3-1) = 6$

The critical value for 6 degrees of freedom and the .01 level of significance is 16.812. The decision rule is to reject the null hypothesis if the computed value of $X^2$  is greater than 16.812

```{r problem16}
AgePresData <- matrix(c(20,18,22,50,46,44,58,63,59,34,43,43), ncol=3, byrow=T)
colnames(AgePresData) <- c('Low', 'Medium', 'High')
rownames(AgePresData) <- c('Less than 25', '25 to 40', '40 to 60', '60+')
print(AgePresData)
chisq.test(AgePresData)
```
We fail to reject the null hypothesis at the 0.01 significance level. The sample data fails to indicate a difference between the Age and amount of pressure are related.

### **Problem 17:**

```{r problem17}
home_with_pool = dplyr::filter(real_estate, Pool==1)
home_without_pool = dplyr::filter(real_estate, Pool==0)
t.test(home_with_pool$Price, home_without_pool$Price, conf.level = .90, var.equal = FALSE)
```

```{r}
home_with_garage = dplyr::filter(real_estate, Garage==1)
home_without_garage = dplyr::filter(real_estate, Garage==0)
t.test(home_with_garage$Price, home_without_garage$Price, conf.level = .90, var.equal = FALSE)
```

```{r}
HMDEFAULT = dplyr::filter(real_estate, Default==1)
HMNONDEFAULT = dplyr::filter(real_estate, Default==0)
t.test(HMDEFAULT$Price, HMNONDEFAULT$Price, conf.level = .95, var.equal = FALSE)
```

```{r problem17_d}
MPrice = median(real_estate$Price);MPrice
PcHighPrice = dplyr::filter(real_estate, Price>=MPrice, Pool==1)
PcLowPrice = dplyr::filter(real_estate, Price<MPrice, Pool==1)
n = nrow(PcHighPrice) + nrow(PcLowPrice)
nsamp = c(n,n)
poolProp = c(nrow(PcHighPrice), nrow(PcLowPrice))
prop.test(poolProp, nsamp, conf.level = .95)

```


```{r problem17_e}
library(ggplot2)

ggplot(real_estate, aes(x = Price)) + 
  geom_histogram(data=dplyr::filter(real_estate, Price>=MPrice, Pool==1), 
                 fill="green", alpha=0.2, position = "dodge") +
  geom_histogram(data=dplyr::filter(real_estate, Price<MPrice, Pool==1),
                 fill="red", alpha=0.2, position = "dodge") +
  labs(title="Homes with Pools with High and Low Price")
```


### **Problem 18:**

```{r problem18_1}
colnames(schools)
nrow(schools)

df_small_districts = dplyr::filter(schools, Students<2000)
df_larger_districts = dplyr::filter(schools, Students>=2000)
mean_1 = round(mean(df_small_districts$Salary),2);mean_1
mean_2 = round(mean(df_larger_districts$Salary),2);mean_2
var.test(df_small_districts$Salary, df_larger_districts$Salary)
t.test(df_small_districts$Salary, df_larger_districts$Salary, conf.level = .95, var.equal = TRUE, alternative = "greater")

```



### **Problem 19:**

```{r problem19}
library(dplyr)
TabMaleDBP = dplyr::filter(blsa, Sex=="M")
TabFemaleDBP = dplyr::filter(blsa, Sex=="F")

t.test(TabMaleDBP$DBP, TabFemaleDBP$DBP, conf.level = .99, var.equal = FALSE)

```



```{r problem19_b}
library(dplyr)
TabM = dplyr::filter(blsa, Sex=="M")
TabFemaleDBP = dplyr::filter(blsa, Sex=="F")

t.test(TabMaleDBP$DBP, TabFemaleDBP$DBP, conf.level = .99, var.equal = FALSE)

```

```{r problem19_c}
group = blsa, fill = blsa +
  stat_boxplot(geom = "errorbar", width = 0.5) +
  geom_boxplot() +
  theme(legend.position = "none") +
  labs(title="Blood Pressure of Smokers and Nonsmokers")
```

### **Problem 20:**

```{r problem20_c}
TableHomePool = table(real_estate$Pool, real_estate$Township)
colnames(TableHomePool) <- c('Town 1', 'Town 2', 'Town 3', 'Town 4', 'Town 5')
rownames(TableHomePool) <- c('Without Pool', 'With Pool')
print(TableHomePool)

chisq.test(TableHomePool)
```

```{r problem20_d}
TableHomeGarage = table(real_estate$Garage, real_estate$Township)
colnames(TableHomeGarage) <- c('Town 1', 'Town 2', 'Town 3', 'Town 4', 'Town 5')
rownames(TableHomeGarage) <- c('Without Garage', 'With Garage')
print(TableHomeGarage)

chisq.test(TableHomeGarage)
```


```{r problem20_e}
library(dplyr)

real_estate %>%
  group_by(Township) %>%
  summarise_at(vars(Price), list(Mean_Price = mean, Median_Price = median), na.rm = T)
```

```{r problem20_f}
library(ggplot2)

ggplot(real_estate, aes(x = Price , y = Township, group = Township, fill = Township)) +
  stat_boxplot(geom = "errorbar", width = 0.5) +
  geom_boxplot() +
  theme(legend.position = "none") +
  labs(title="Township and Prices Boxplot")

```


```{r problem20_g}

library(ggplot2)
ggplot(real_estate, aes(x=Price, y=Days)) +
  geom_point() +
  labs(title = "Price and Days", 
       x = "Price", 
       y = "Days")

```
